
\begin{definition}[consistent]
  A set of \wff $\Sigma$ is consistent if $\Sigma\nvdash\alpha$ or $\Sigma\nvdash\neg\alpha$ for all \wff $\alpha$.
\end{definition}

\begin{claim}
  A set of \wff $\Sigma$ is consistent iff $\exists\alpha, \Sigma\nvdash\alpha$.
\end{claim}

\begin{proof}

  $\rightarrow$: pick any $\alpha$, either $\Sigma\nvdash\alpha$ or $\Sigma\vdash\alpha$, then there exists one that is not proved by $\Sigma$.

$\leftarrow$: assume that for some $beta$, $\Sigma\nvdash\beta$, if the first definition is violated, then, for some $\alpha$, $\Sigma\vdash\alpha$ and $\Sigma\vdash\neg\alpha$, then $\Sigma\vdash\Sigma\cup\{\alpha,\neg\alpha\}\vdash\{\alpha,\neg\alpha\}$. $\{\alpha, \neg\alpha\}\vdash\beta$ for every $\beta$, contradicting the assumption.
\end{proof}

\begin{corollary}
If $\Sigma\subseteq\Sigma'$, then if $\Sigma$ is consistent, then so does $\Sigma'$.
\end{corollary}

\begin{definition}[Soundness]
If $\vdash\alpha$, then $\alpha$ is a tautology.
\end{definition}

\begin{corollary}
$\nvdash P$, since $P$ is a propositional variable, which is not a tautology.
\end{corollary}

The universe $U$ (the set which includes everything) is the least consistent, since it contains everything. Therefore, for every $\alpha\in U$, $\neg\alpha\in U$, $U$ can prove both. It has the most unsteady stance:).

$\vdash$ is about syntax while $\vDash$ is about semantics. $\vdash$ shows that though the machine doesn't know anything, it can reach some result by axioms and modus ponens. $\vDash$ shows that the result is true in the real world given by human being's assignment of truth values.

\begin{theorem}
In a sound proof system, every satisfiable $\Sigma$ is consistent. 
\end{theorem}

\begin{proof}
\bwoc, assume the satisfiable $\Sigma$ is inconsistent. Then, for some $\alpha$, both $\Sigma\vdash\alpha$ and $\Sigma\vdash\neg\alpha$. 
If $\Sigma$ is satisfiable, then for some truth assignment $V$, $V$ satisfies all \wff in $\Sigma$. By soundness, $\Sigma\vDash\alpha$ and $\Sigma\vDash\neg\alpha$. So for that assignment $V$, we get $V(\alpha)=T$ and $V(\neg\alpha)=T$, violating the truth table of negation.
\end{proof}

The above theorem builds a bridge from semantic to syntax.

\begin{theorem}[Extended Soundness]
If $Sigma\vdash\alpha$, then $\Sigma\vDash\alpha$.
\end{theorem}

\begin{corollary}
If any set of \wff is consistent, then, in particular $\emptyset$.
\end{corollary}


\begin{definition}[Maximally Consistent]
We say that $\Sigma$ is maximally consistent if $\Sigma$ is consistent, but, for every $\alpha$, either $\Sigma\vdash\alpha$ or $\Sigma\cup\{\alpha\}$ is inconsistent.
\end{definition}

Simialr to the definition of maximally satisfiable, the system has the maximal knowledge, with which it can prove everything and rejects to embrace any unprovable things cause it will cause contradiction.

\begin{mdframed}
\begin{example}
Let $\Sigma\equiv\{P_1\}$ over the variables $P_1, P_2, \ldots$.
\begin{claim}
$\{P_1\}$ is consistent since it's satisfiable.
\end{claim}
\begin{claim}
$\{P_1\}$ is not maximally consistent
\end{claim}
\begin{proof}
  It suffices to show that $\{P_1\}\nvdash P_3$ and $\{P_1, P_3\}$ is consistent.

  By soundness, it suffices to show that $\{P_1\}\vDash P_3$. Just find some truth assignment that $V$ \st $V(P_1)=T$ and $V(P_3)=F$. So $\{P_1\}\nvdash P_3$.
  
  Since $\{P_1, P_3\}$ is satisfiable, so it's consistent.
\end{proof}
\end{example}
\end{mdframed}

\begin{lemma}
For every consistent $\Sigma$, there exists a maximally consistent $\Sigma' \supseteq \Sigma$.
\end{lemma}

\begin{proof}
Let $\{\alpha_1, \ldots, \alpha_i, \alpha_{i+1}, \ldots\}$ is a set of \wff over $\{P_1,P_2,\ldots,P_i,\ldots\}$
\end{proof}

\begin{definition}[Monotonicity]
  $\forall \Sigma, \Sigma', \alpha$, if $\Sigma\vdash\alpha, \Sigma\subseteq\Sigma'$, then $\Sigma'\vdash\alpha$.
\end{definition}