\begin{theorem}[Deduction Theorem]
  $\Sigma \cup \{ \alpha \} \vdash \beta$ iff $\Sigma \vdash \alpha \rightarrow \beta$.
\end{theorem}


\begin{theorem}
  Any consistent set of \wff $\Sigma$ can be extended to $\Sigma$' \st $\Sigma' \supseteq \Sigma$ that is maximally consistent.
\end{theorem}


\begin{proof}
We construct a sequence of set of \wff. $\Sigma_0 \subseteq \Sigma_1 \subseteq \Sigma_2 \subseteq \ldots \subseteq \Sigma_i \subseteq \Sigma_{i+1}$ \st 1) $\Sigma_0 = \Sigma$ and 2) For all $\Sigma_i$, $\Sigma_i$ is consistent, and 3) for fixed enumeration of all \wff., $\alpha_1, \alpha_2, \ldots, \alpha_n$, for all $i$, either $\Sigma_i \vdash \alpha$ or $\Sigma_i \vdash \neg \alpha$.

Assume $\Sigma_i$ is defined and meets the requirements, let $\Sigma_{i+1}$ be $\Sigma_i$ if $\Sigma_i \vdash \neg \alpha_{i}$, otherwise $\Sigma_{i+1} = \Sigma_i \cup \{ \alpha_i \}$. 

Now we want to prove if $\Sigma_i$ meets the requirement, then so will $\Sigma_{i+1}$.
Requirement 1) is trivial. Requirement 2) and 3) are simultaneously proved by showing $\Sigma_{i+1} \nvdash \neg \alpha_i$ if $\Sigma_{i+1} = \Sigma_i \cup \{\alpha_i\}$, which is shown by the following claim.

\begin{claim}
  $\forall \Sigma, \alpha$, if $\Sigma \cup \{ \alpha \} \vdash \neg \alpha$, then $\Sigma \vdash \neg \alpha$. 
\end{claim}
\begin{proof}
  To prove this claim, it suffices to show $\vdash (\alpha \rightarrow \neg \alpha) \rightarrow \neg \alpha$ based on deduction theorem, which is a tautology.
\end{proof}

Now we show the $\Sigma'=\cup_{i=1}^\infty \Sigma_i$ is maximally consistent.
To prove it, it suffices to show that $\forall \alpha_i, i \in N$, $\Sigma' \vdash \alpha$ or $\Sigma' \vdash \neg \alpha$.
By requirement 2), $\Sigma_i \vdash \alpha_i$ or $\Sigma_i \vdash \neg \alpha_i$. Now we are going to show $\Sigma' \vdash \alpha_i$ or $\Sigma' \vdash \neg \alpha_i$.

\bwoc, suppose $\Sigma'$ is inconsistent. In this case, for some $\alpha$, $\Sigma'\vdash\alpha\ and\ \Sigma'\vdash\neg\alpha$. Let $\beta_1,\ldots\beta_k$ be the proof of $\alpha$ from $\Sigma'$ and $\gamma_1,\ldots,\gamma_l$ be the proof of $\neg\alpha$ from $\Sigma'$.

Each $\beta_i$ that is an assumption from $\Sigma'$ belongs to some $\Sigma_{m_i}$, and each $\gamma_i$ that is an assumption from $\Sigma'$ belongs to some $\Sigma_{m_j}$.
Since both formal proofs of $\alpha$ and of $\neg\alpha$ are finite, there is some $i^*$ that is bigger than all of these $m_i$'s and $m_j$'s. Therefore, for each $beta_i$ or $gemma_j$ that are used as assumptions, $\beta_i, \gamma_j \in \Sigma_{i^*}$. Now by monotonicity, $\Sigma_{i^*} \vdash \alpha$ and $\Sigma_{i^*} \vdash \neg \alpha$, which is a contradiction to requirement 2).

\end{proof}

\begin{theorem}
  In a sound proof system, every consistent $\Sigma$ is satisfiable.
\end{theorem}


\begin{proof}
Let $\Sigma'$ be a maximally consistent set of \wff.s \st $\Sigma\subseteq\Sigma'$. Define a truth assignment $V_{\Sigma'}$ as follows: $V_{\Sigma'}(P_i) = T$ iff $\Sigma' \vdash P_i$ where $P_i$ is a propositional variable.
Now extend it into  $V_{\Sigma'}(\alpha_i) = T$ iff $\Sigma' \vdash \alpha_i$.

\end{proof}


\begin{proof}
$\Sigma$ is satisfiable by all-truth assignemnt, so it's consistent.
To show it's maximally consistent, we need to show for every $\alpha$, if $\Sigma\nvdash\alpha$, then $\Sigma\cup\{\alpha\}$ is inconsistent.
If $\Sigma\nvdash\alpha$, then $\Sigma\nvDash\alpha$, then $\Sigma\cup\{\alpha\}$ is unsatisfiable, so it's inconsistent.

\end{proof}